Windowed Developer Topic Analysis

### 2009-03-16 lda-paper.tex
lda paper

Topic analysis, with respect to Software Control Systems (SCSs), is
valuable as it has many uses for the stakeholders involved.  Example
scenarios where the usage of these techniques would prove useful are
numerous. Imagine a development team has agreed to implement some
features at the start of an iteration. By the end of the iteration
only some of the agreed upon features are implemented. The developers
swear they were focusing solely on the features agreed upon. Topic
analysis could help by teasing out the distinct topics or focusses of
the developers. Perhaps the developers had to deal with issues
orthogonal to the features they were working on, but were probably
necessary to complete such features. Developers might have grouped
those orthogonal changes into one of the features they were working
on, thus when asked about how their time was spent the orthogonal
changes would not be covered. This is what we hope that windowed topic
analysis can uncover. If windowed topic analysis can discover
important but overlooked topics, it can help both developers and
managers.

We propose to extend common concept analysis techniques by executing
them over windows of documents over time. We hope by looking at topic
clusters over time we can I identify continuous effort on a select
group of topics.

Topic analysis of fine-grained revisions is important as often
revisions belonging to different topics are interleaved or even mixed
inside of commits.  Sometimes a single revision to a file contains
multiple bug fixes.  Using powerful topic analysis like Latent
Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can
seperate these mixed signals into coherent topics. In this case a
topic is a set of tokens that are related to each other.

@@

Topic analysis is useful because it can help us partition a project's
time-line into sub-iterations. By breaking apart an iteration we might
be able to extract the underlying software development process from
these artifacts.

In this paper we seek to explore how topics shift over time in
the source control repository of a project. We analyze commit
comments, and even source code per revision per each time window and
extract the topics in that time window. We expect that topic sets
will change over time and the change in topic sets indicate a change
in the focus of development.

@@

Latent Dirichlet Alloation (LDA)~\cite{944937} is a competitor to
latent semantic indexing (LSI), probabilistic latent semantic indexing
(pLSI) and semantic clustering. It is used for document modelling,
document clustering and collaborative filtering. What LDA attempts to
do is model a documents as mixture model of topics.  The common use of
LDA in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

@@

@@

The document we are primarily using are the changelog comments, these
are the comments added when a revision to the source code is
committed, it could be per file or per commit of changes to multiple
files, this depends on the source control system used.

Documents are converted into word count distributions. The
distributions include all of the possible words in the system.

@@

We then create windows which are sets of documents grouped by time
window. Time windows can overlap but don't have to. Windowing by time
allows for many levels of granularity. We usually used one month as
the length of our time windows.

%   4. [ ] Cluster Windowed Set w/ LDA
\subsection{Cluster Windowed Set w/ LDA}
%   5. [ ] Extract Clusters

We take our topic clusterer and cluster the documents by their topics
and get the $n$ unique topic clusters out. In this paper we used LDA
but LSI could be used in a similar way.

This is a very slow step as LDA takes a long time to run.

\subsection{Extract Clusters}

@@

Once we get our clusters, we analyze them and look for repeating
topics. We do this by taking the top 10 most common words and
comparing them. Given a fraction such as 8/10 we joined topics where
8/10 of the top ten topic words were shared.

Clusters in the same period which are this similar are grouped
together. Anything that matches either of those clusters 8/10 will
match.

\subsection{Continuous Cluster Analysis}

We extend the cluster matching across contiguous months.q

%   8. [ ] Topic Smear
\subsection{Topic Smear}
%   9. [ ] Visualization [0/1]
\subsection{Visualization}
%       1. [ ] need the visualizer

@@

@@

Our analysis tools consisted of: Hiraldo-Grok, an OCaml based spin off
of Grok used for answering queries; Gnuplot, a graph plotting package,
lda-c a LDA package by Biel

@@

A subsampling of the notable topic words is displayed in Table
\ref{tab:portability}. We tracked one topic cluster which seemed
related to portability.

@@

%3. [ ] Previous Work [0/2]
%   2. [ ] LDA Work [0/6]
\section{Results}
%   1. Interesting smears 
%   2.   postgresql
\subsection{Postgresql}
%   3.   mysql
\subsection{Mysql}
%   4.   firebird
\subsection{Firebird}
%   5.   what's that other db?
\subsection{ what's that other db?}
%7.   Validation [0/1]
\subsection{Validation}
%   1. well i'm stumped
%   2. look at it?
%   3. investigate those revisions and check the proportion etc
%   4. exploratory
%   5.   work it out
%8.   Validity Threats [0/5]
\section{Validity Threats}

@@

>>>>

Topic analysis, with respect to Software Control Systems (SCSs), is
valuable as it has many uses for the stakeholders involved.  Example
scenarios where the usage of these techniques would prove useful are
numerous. Imagine a development team has agreed to implement some
features at the start of an iteration. By the end of the iteration
only some of the agreed upon features are implemented. The developers
swear they were focusing solely on the features agreed upon. Topic
analysis could help by teasing out the distinct topics or focusses of
the developers. Perhaps the developers had to deal with issues
orthogonal to the features they were working on, but were probably
necessary to complete such features. Developers might have grouped
those orthogonal changes into one of the features they were working
on, thus when asked about how their time was spent the orthogonal
changes would not be covered. This is what we hope that windowed topic
analysis can uncover. If windowed topic analysis can discover
important but overlooked topics, it can help both developers and
managers. Developer can reflect and see a good summary of what worked
on, and managers can see the focus of development while still
remaining relatively high level.

We propose to extend common concept analysis techniques by executing
them over windows of documents over time. We hope by looking at topic
clusters over time we can identify continuous effort on a select
group of topics.

Topic analysis of fine-grained revisions is important as often
revisions belonging to different topics are interleaved or even mixed
inside of commits.  Sometimes a single revision to a file contains
multiple bug fixes.  Using powerful topic analysis techniques like
Latent Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA)
one can seperate these mixed signals into coherent topics. In this
case a topic is a set of tokens that are related to each other.

@@

Another use of topic analysis is to help us partition a project's
time-line into sub-iterations of developer focus. By breaking apart an
iteration we might be able to extract the underlying software
development process from these artifacts or at the very least what
maintenance was occurring in the repository at that time.

In this paper we seek to explore how topics shift over time in the
source control repository of a project. We analyze commit comments per
revision in each time window and we extract the topics in that time
window. We expect that topics in each window will change over time and
the similarities and differences in these sets of topics will indicate
both developer focus and changes in developer focus.

@@

Latent Dirichlet Alloation (LDA)~\cite{944937} is a competitor to
Latent Semantic Indexing (LSI), probabilistic Latent Semantic Indexing
(pLSI) and semantic clustering. It is used for document modelling,
document clustering and collaborative filtering. What LDA attempts to
do is model a documents as mixture model of topics.  The common use of
LDA in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

@@

A subsampling of the notable topic words is displayed in Table
\ref{tab:portability}. We tracked one topic cluster which seemed
related to portability. We picked out terms which seem to summarize
those topics.

@@

The document we are analyzing are the changelog comments, these are
the comments added when a revision to the source code is committed.
Documents are converted into word count distributions. These
distributions are can be normalized by by the size of the
message. After all messages are processes the distributions are
extended to include all words in all distributions (each distribution
is extended with word bins of count 0).

@@

Given all messages, we group the messages into windows. We could use
overlapping windows, but in this paper we use non-overlapping windows
of a set time unit.  Windowing by time allows for many levels of
granularity. We usually used one month as the length of our time
windows, while we could use different granularities for the study we
think that a month is a sizable unit of development (4 weeks of
development).

\subsection{Apply Topic Analysis to each Window}

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topic clusters, we used $20$
topics in our case study. Our Topic Analysis tool is an implementation
of LDA, but we could've used other similar tools like LSI.

This is a very slow step as LDA takes a long time to run, and takes
even longer since we run it once per window.

%\subsection{Extract Clusters}

@@

Once we have our topics extracted for each window , we analyze them
and look for repeating topics. We do this by taking the top 10 most
common words and then comparing them.

We then abstract these topics by comparing them to each other.  Each
topic is compared to each other topic in the system, given a certain
threshold of top 10 similarity, like 8/10 words. Then given this topic
similarity matrix, we find the transitive closure of similar topics,
that is we fill flood along the similarity arcs until we have
partitioned the topics into clusters of similar topics.  These
clusters of topics are called trends. A trend is probably interesting
and valuable if there is more than 1 topic in the trend.

This technique has weaknesses in that nodes a few neighbors away in
similarity might not share any similar words.  We do this because we
want to be able to color or highlight topics which are related to each
other and track them throughout time.

Once we have determined our similarity clusters we can choose to
analyze and to plot the topics.

\begin{comment}
Topic Clustering
 · Need to track continuous topics across
 · Similarity between topics
 · Clusters of the transitive closure of topics with X%
   similarity
    ­ Fill flood along similarity, make subsets of
      everyone who X% similar to any of their neighbors,
      make that a cluster
\end{comment}

\subsection{Visualization}
%       1. [ ] need the visualizer

We have come up with multiple ways to visualize these topics and
trends.  For all of these plots if we trends which have continuous
segments we plot those segments as joined across time.  We have a
compressed stacked trend view, which displays trends arbitrarily
across the y axis but are placed where they occur on the time-based x
axis. Another plot we have is the trend-time plot, a each trend gets
it own distinct-y value , while the x-axis is time; these topics are
then each plotted on their own line across time as they occur. The
last plot we have is the histogram plot, where we plot each trend on
its own line but stack up the segments of the trend, much like a plot
of counts or a histogram.

The first visualization is a compressed view, we try to sink the
larger trends to the bottom, once the larger trends are stacked, we
fill in the gaps with the smaller trends in the same window and then
stack the smaller trends on top. Although there is a chance that gaps
will be left due to nonoptimal stacking in practice there are lot of
small trends (90\% of all trends contain 1 topic) which fill in these
gaps quickly. The trend segments that share the same color are the
same trend, colors are randomly chosen for each trend, so color
similarity is meaningless, only exact color matching matters.
The compressed view is nice because it shows all of the
information and the repeating continuous trends are easy to pick out,
unfortunately repeating discontinuous trends are harder to spot.

The second visualization is the trend-time plot, it displays repeating
trends more clearly by dedicating a horizonal line for trend segments
dedicated to one trend. This means that if a trend contains
discontinuous segments the segments appear on the same line. One down
side to this view is the least common trends need to be pruned
otherwise it'd take too many horizontal lines to plot them.

The 3rd visualization is the histogram plot which is a lot like the
time-trend plot except the trends are plotted together by stacking to
their row so time information is lost and the trends are ordered by
the number of topics in the trend. The histogram plot shows the count
of instances of a trend and thus indicates which trends occur the
most. Unfortunately due to the large number of topics given allotted
space it is often best to crop off the trends with only one topic,
otherwise the tail is long.

@@

We extracted the repositories of: MySQL, PostGreSQL, MaxDB and
Firebird.

@@

Our analysis tools consisted of: Hiraldo-Grok, an OCaml based spin off
of Grok used for answering queries; Gnuplot, a graph plotting package;
lda-c, a LDA package implemented by Biel; and our topic plotter,
implemented in Haskell.

@@

@@

%3. [ ] Previous Work [0/2]
%   2. [ ] LDA Work [0/6]

\section{Results}
%   1. Interesting smears 
%   2.   postgresql
\subsection{Postgresql}
%   3.   mysql
\subsection{Mysql}
%   4.   firebird
\subsection{Firebird}
%   5.   what's that other db?
\subsection{ what's that other db?}
%7.   Validation [0/1]

\subsection{Compare to topics over entire development}

\subsection{Validation}

%   1. well i'm stumped
%   2. look at it?
%   3. investigate those revisions and check the proportion etc
%   4. exploratory
%   5.   work it out
%8.   Validity Threats [0/5]
\section{Validity Threats}

@@

We truncated the topics to top 10 tokens, this summary of a topic
might not have been as useful as determining the actual topic distance
between two word topic distributions.
